---
output:
  html_document:
    theme: yeti  # many options for theme, this one is my favorite.
params:
  threads: 20
  rdata: ""
  token: ""
  ref: ""
  out: ""
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include=FALSE}
#load the data from the previous chapter as needed
current_chapter = 13
previous_chapter = current_chapter-1
load(file = paste0("./chapter_images/chapter",previous_chapter,".RData"))
```

# Identifying Marker Peaks

We have introduced the identification of marker features previously in the chapter discussing gene scores. The same function (`getMarkerFeatures()`) can be used to identify marker features from any matrix stored in an `ArchRProject`. Marker features are features that are unique to a specific cell grouping. These can be very useful in understanding cluster- or cell type-specific biology. In this chapter, we will discuss how to use this function to identify marker peaks.

## Identifying Marker Peaks with ArchR

Often times, we are interested to know which peaks are unique to an individual cluster or a small group of clusters. We can do this in an unsupervised fashion in ArchR using the `addMarkerFeatures()` function in combination with `useMatrix = "PeakMatrix"`.

First, lets remind ourselves of the cell types that we are working with in `projHeme5` and their relative proportions.
```{r, collapse=TRUE}
table(projHeme5$Clusters2)
```

Now, we are ready to identify marker peaks by calling the `addMarkerFeatures()` function with `useMatrix = "PeakMatrix"`. Additionally, we tell ArchR to account for differences in data quality amongst the cell groups by setting the `bias` parameter to account for TSS enrichment and the number of unique fragments per cell.

```{r, collapse=TRUE}
markerPeaks <- getMarkerFeatures(
	ArchRProj = projHeme5, 
	useMatrix = "PeakMatrix", 
	groupBy = "Clusters2",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  testMethod = "wilcoxon"
)
```

The object returned by the `getMarkerFeatures()` function is a `SummarizedExperiment` that contains a few different `assays`.
```{r, collapse=TRUE}
markerPeaks
```

We can use the `getMarkers()` function to retrieve particular slices of this `SummarizedExperiment` that we are interested in. The default behavior of this function is to return a list of `DataFrame` objects, one for each cell group.
```{r, collapse=TRUE}
markerList <- getMarkers(markerPeaks, cutOff = "FDR <= 0.01 & Log2FC >= 1")
markerList
```

If we are interested in the marker peaks for a specific cell group, we can access this from the list via the `$` accessor.
```{r, collapse=TRUE}
markerList$Erythroid
```

Instead of a list of `DataFrame` objects, we can use `getMarkers()` to return a `GRangesList` object by setting `returnGR = TRUE`.
```{r, collapse=TRUE}
markerList <- getMarkers(markerPeaks, cutOff = "FDR <= 0.01 & Log2FC >= 1", returnGR = TRUE)
markerList
```


This `GRangesList` object can similarly be subset to a `GRanges` object for a particular cell group using the `$` accessor.
```{r, collapse=TRUE}
markerList$Erythroid
```

## Plotting Marker Peaks in ArchR

ArchR provides multiple plotting functions to interact with the `SummarizedExperiment` objects returned by `getMarkerFeatures()`.

### Marker Peak Heatmaps

We can visualize these marker peaks (or any features output by `getMarkerFeatures()`) as a heatmap using the `markerHeatmap()` function.

```{r, collapse=TRUE}
heatmapPeaks <- markerHeatmap(
  seMarker = markerPeaks, 
  cutOff = "FDR <= 0.1 & Log2FC >= 0.5",
  transpose = TRUE
)
```

We can plot this heatmap using `draw()`.
```{r, collapse=TRUE}
draw(heatmapPeaks, heatmap_legend_side = "bot", annotation_legend_side = "bot")
```

To save an editable vectorized version of this plot, we use the `plotPDF()` function.
```{r, collapse=TRUE}
plotPDF(heatmapPeaks, name = "Peak-Marker-Heatmap", width = 8, height = 6, ArchRProj = projHeme5, addDOC = FALSE)
```

### Marker Peak MA and Volcano Plots

Instead of plotting a heatmap, we can also plot an MA or Volcano plot for any individual cell group. To do this, we use the `plotMarkers()` function. For an MA plot we specify `plotAs = "MA"`. Here we specify the "Erythroid" cell group via the `name` parameter.
```{r, collapse=TRUE}
pma <- plotMarkers(seMarker = markerPeaks, name = "Erythroid", cutOff = "FDR <= 0.1 & Log2FC >= 1", plotAs = "MA")
pma
```

Similarly, for a Volcano plot, we specify `plotAs = "Volcano"`.
```{r, collapse=TRUE}
pv <- plotMarkers(seMarker = markerPeaks, name = "Erythroid", cutOff = "FDR <= 0.1 & Log2FC >= 1", plotAs = "Volcano")
pv
```

To save an editable vectorized version of these plots, we use the `plotPDF()` function.

```{r, collapse=TRUE}
plotPDF(pma, pv, name = "Erythroid-Markers-MA-Volcano", width = 5, height = 5, ArchRProj = projHeme5, addDOC = FALSE)
```

### Marker Peaks in Browser Tracks

Additionally we can see these peak regions overlayed on our browser tracks by passing the relevant peak regions to the `features` parameterin the `plotBrowserTrack()` function. This will add an additional BED-style track of marker peak regions to the bottom of our ArchR track plot. Here we specify plotting the GATA1 gene via the `geneSymbol` parameter.

```{r, collapse=TRUE}
p <- plotBrowserTrack(
    ArchRProj = projHeme5, 
    groupBy = "Clusters2", 
    geneSymbol = c("GATA1"),
    features =  getMarkers(markerPeaks, cutOff = "FDR <= 0.1 & Log2FC >= 1", returnGR = TRUE)["Erythroid"],
    upstream = 50000,
    downstream = 50000
)
```

We can plot this using `grid::grid.draw()`.
```{r, collapse=TRUE}
grid::grid.newpage()
grid::grid.draw(p$GATA1)
```

To save an editable vectorized version of this plot, we use the `plotPDF()` function.

```{r, collapse=TRUE}
plotPDF(p, name = "Plot-Tracks-With-Features", width = 5, height = 5, ArchRProj = projHeme5, addDOC = FALSE)
```

## Pairwise Testing Between Groups

Marker feature identification is a very specific type of differential test. However, ArchR also enables standard differential testing using the same `getMarkerFeatures()` function. The trick is to set `useGroups` to one of the two cell groups and `bgdGroups` to the other cell group. This performs a differential test between the two provided groups. To be explicitly clear, the cell groups here are user-defined. They can be anything that fits your application. All you need to do is add a new column to `cellColData` that defines the groups that you would like to compare. In all differential tests, the peaks that are higher in the group passed to `useGroups` will have positive fold change values while the peaks that are higher in the group passed to `bgdGroups` will have negative fold change values. 

Here we perform a pairwise test between the "Erythroid" cell group and the "Progenitor" cell group.

```{r, collapse=TRUE}
markerTest <- getMarkerFeatures(
  ArchRProj = projHeme5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters2",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Erythroid",
  bgdGroups = "Progenitor"
)
```

We can then plot an MA or Volcano plot using the `plotMarkers()` function. We indicate an MA plot using `plotAs = "MA"`.

```{r, collapse=TRUE}
pma <- plotMarkers(seMarker = markerTest, name = "Erythroid", cutOff = "FDR <= 0.1 & abs(Log2FC) >= 1", plotAs = "MA")
pma
```

Similarly, we plot a volvano plot using `plotAs = "Volcano"`.
```{r, collapse=TRUE}
pv <- plotMarkers(seMarker = markerTest, name = "Erythroid", cutOff = "FDR <= 0.1 & abs(Log2FC) >= 1", plotAs = "Volcano")
pv
```

To save an editable vectorized version of these plots, we use the `plotPDF()` function.

```{r, collapse=TRUE}
plotPDF(pma, pv, name = "Erythroid-vs-Progenitor-Markers-MA-Volcano", width = 5, height = 5, ArchRProj = projHeme5, addDOC = FALSE)
```

We will continue this differential analysis in the next chapter by looking for motif enrichments in our differentially accessible peaks.


```{r saveImage, include=FALSE}
filename <- paste0("./chapter_images/chapter",current_chapter,".RData")
remove(current_chapter)
remove(previous_chapter)
save.image(file =  filename)
```