---
output:
  html_document:
    theme: yeti  # many options for theme, this one is my favorite.
params:
  threads: 20
  rdata: ""
  token: ""
  ref: ""
  out: ""
---

```{r, include=FALSE, eval=FALSE}
params2 <- params
unlockBinding("params", env = .GlobalEnv)
#load("Chapter-13-Footprints.Rdata")
load(params$rdata)
params <- params2
rm(params2)
devtools::install_github("GreenleafLab/ArchR", 
  auth_token = params$token, 
  ref = params$ref,
  repos = BiocManager::repositories(),
  dependencies = FALSE
)
library(ArchR)
fn <- unclass(lsf.str(envir = asNamespace("ArchR"), all = TRUE))
fn <- fn[!grepl("\\.", fn)]
fn <- fn[fn!="ArchRProj"]
for (i in seq_along(fn)){
    tryCatch({
        eval(parse(text = paste0(fn[i], "<-ArchR::", fn[i])))
    }, error = function(x) {
    })
}
addArchRThreads(threads = params$threads)
addArchRGenome("hg19")
# fn <- unclass(lsf.str(envir = asNamespace("ArchR"), all = TRUE))
# fn <- fn[fn!="ArchRProj"]
# for (i in seq_along(fn)) {
#     tryCatch({
#         eval(parse(text = paste0(fn[i], "<-ArchR:::", fn[i])))
#     }, error = function(x) {
#     })
# }
set.seed(1)
```

# Session Information
This tutorial was run on the date specified below.
```{r eval=FALSE}
Sys.Date()
```
> \## [1] "2020-04-15"  

The `sessionInfo()` at run time was:
```{r eval=FALSE}
sessionInfo()
```
> \## R version 3.6.1 (2019-07-05)  
## Platform: x86_64-pc-linux-gnu (64-bit)  
## Running under: CentOS Linux 7 (Core)  
##   
## Matrix products: default  
## BLAS/LAPACK: /share/software/user/open/openblas/0.2.19/lib/libopenblasp-r0.2.19.so  
##   
## locale:  
##  [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
##  [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
##  [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
##  [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
## [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       
##   
## attached base packages:  
##  [1] grid      parallel  stats4    stats     graphics  grDevices utils    
##  [8] datasets  methods   base     
##   
## other attached packages:  
##  [1] circlize_0.4.8              ComplexHeatmap_2.2.0       
##  [3] gridExtra_2.3               pdftools_2.3               
##  [5] gtable_0.3.0                ggrastr_0.1.7              
##  [7] ArchR_0.9.1                 magrittr_1.5               
##  [9] rhdf5_2.30.1                Matrix_1.2-17              
## [11] data.table_1.12.8           SummarizedExperiment_1.16.1  
## [13] DelayedArray_0.12.2         BiocParallel_1.20.1        
## [15] matrixStats_0.56.0          Biobase_2.46.0             
## [17] GenomicRanges_1.38.0        GenomeInfoDb_1.22.1        
## [19] IRanges_2.20.2              S4Vectors_0.24.3           
## [21] BiocGenerics_0.32.0         ggplot2_3.2.1              
##   
## loaded via a namespace (and not attached):  
##  [1] nlme_3.1-140                      bitops_1.0-6                     
##  [3] fs_1.3.1                          usethis_1.5.1                    
##  [5] devtools_2.2.1                    RColorBrewer_1.1-2               
##  [7] rprojroot_1.3-2                   tools_3.6.1                      
##  [9] backports_1.1.6                   R6_2.4.1                         
## [11] mgcv_1.8-28                       lazyeval_0.2.2                   
## [13] colorspace_1.4-1                  GetoptLong_0.1.8                 
## [15] withr_2.1.2                       tidyselect_1.0.0                 
## [17] prettyunits_1.1.1                 processx_3.4.1                   
## [19] curl_4.3                          compiler_3.6.1                   
## [21] cli_2.0.1                         Cairo_1.5-10                     
## [23] desc_1.2.0                        labeling_0.3                     
## [25] rtracklayer_1.46.0                scales_1.1.0                     
## [27] hexbin_1.28.1                     askpass_1.1                      
## [29] callr_3.4.1                       nabor_0.5.0                      
## [31] stringr_1.4.0                     digest_0.6.25                    
## [33] Rsamtools_2.2.3                   rmarkdown_2.1                    
## [35] BSgenome.Hsapiens.UCSC.hg19_1.4.0 XVector_0.26.0                   
## [37] pkgconfig_2.0.3                   htmltools_0.4.0                  
## [39] sessioninfo_1.1.1                 BSgenome_1.54.0                  
## [41] rlang_0.4.5                       GlobalOptions_0.1.1              
## [43] farver_2.0.3                      shape_1.4.4                      
## [45] gtools_3.8.2                      dplyr_0.8.4                      
## [47] RCurl_1.98-1.1                    GenomeInfoDbData_1.2.2           
## [49] Rcpp_1.0.4                        munsell_0.5.0                    
## [51] Rhdf5lib_1.8.0                    fansi_0.4.1                      
## [53] lifecycle_0.1.0                   stringi_1.4.6                    
## [55] yaml_2.2.1                        zlibbioc_1.32.0                  
## [57] pkgbuild_1.0.6                    ggrepel_0.8.1                    
## [59] crayon_1.3.4                      lattice_0.20-38                  
## [61] splines_3.6.1                     Biostrings_2.54.0                
## [63] knitr_1.27                        ps_1.3.2                         
## [65] pillar_1.4.3                      rjson_0.2.20                     
## [67] codetools_0.2-16                  pkgload_1.0.2                    
## [69] XML_3.99-0.3                      glue_1.4.0                       
## [71] evaluate_0.14                     qpdf_1.1                         
## [73] remotes_2.1.0                     png_0.1-7                        
## [75] testthat_2.3.1                    purrr_0.3.3                      
## [77] clue_0.3-57                       assertthat_0.2.1                 
## [79] xfun_0.12                         tibble_2.1.3                     
## [81] GenomicAlignments_1.22.1          memoise_1.1.0                    
## [83] cluster_2.1.0                     ellipsis_0.3.0  
